### 2019-02-02 Seneca.tex
mz - Add background, Seneca, related_work

To facilitate model optimization on serverless architecture, we have developed a cloud service, namely Seneca. The Seneca service pipeline consists of

>>>>

To facilitate model optimization on serverless architecture, we have developed Seneca, a framework of  hyperparameter tuning for machine learning algorithm on AWS Lambda. The automated Seneca pipeline consists of packaging, deployment, serverless function optimization and hyperparameter tuning. Figure~\ref{fig:seneca} portrays the architecture of Seneca: the upper right front are three inputs from users: configuration file, specified model, and lambda function. Configuration file is where users specify all hyperparameter options and used by Seneca to create search space. Specified model is the implementation library of training and testing for certain machine learning model and provides clearly defined methods. Lambda function is the function handler that executes on AWS lambda. It downloads the dataset from S3 bucket, splits it by desired training and testing ratio, trains the specified model and returns the testing score.

\subsection{Packaging and Deployment}

After receiving these three inputs, Seneca launches a docker container that simulates the runtime of AWS Lambda. The container will compress the lambda function with all required libraries from specified model in a zip package and upload to S3 bucket. Using S3 bucket as an intermediate storage is to avoid the size limitation (10MB) of deployment package directly transported to AWS Lambda. Once uploaded to S3 bucket, docker container will deploy the deployment package to AWS lambda by AWS Command line Interface~\cite{ref:awscli} for Lambda.

\subsection{Allocated memory optimization}

Seneca has a component of optimizer that allows user to optimize allocated memory of lambda function. The mechanism of optimizer is based on our hypothesis of disproportional increase of computational power bound to allocated memory of lambda function. In other words, additional computational power brought by extra allocated memory reduces the billed duration and possibly overall cost as well. Seneca optimizer tries to find this "sweet-spot" of allocated memory leading to lowest compute charge.

Instead of exhaustively searching for all 46 possible memory configurations of lambda function, we designed a heuristic that complete the task faster and more economically. The detailed heuristic is outlined in the Algorithm~\ref{algo:optimizer}. Seneca optimizer first probes function by a user-defined payload to obtain the memory used by function as start point. It also defines two double-end queues (deque) to store a certain number (N) of historical allocated memory and compute charge data. While the current allocated memory is less than or equal to 3008 MB, optimizer keeps probing function and calculated compute charge by current allocated memory and billed duration. Two exit conditions are established in the heuristic: first, compute charge monotonically increases in the whole length of deque; second, the increase slope is greater than a threshold. Once optimizer find both conditions hold, it will pop the left-most value from allocated memory deque and configure the lambda function at such memory. After repetitive experiments, we find the setting of N = 5 and slope = 1 has the highest possibility to find the lowest compute charge.

\begin{algorithm}[]
\SetAlgoLined
\KwData{Typical payload}
\KwResult{Optimal allocated memory}
Find memory used by payload as start point\;
Define deques for struct of allocated memory \& compute charge\;
 \While{allocated memory $\leq$ 3008 MB}{

  \eIf{compute charge monotonically increase in deque \& slope $\geq$ 1}{
   Popleft from deque\;
   Configure allocated memory of function\; 
   Exit\;
   }{
   Increase allocated memory by 64 MB\;
   Probe lambda function\;
   Append new struct to deque\;
  }
 }
 \caption{Seneca Optimizer Heuristic}
 \label{algo:optimizer}
\end{algorithm}

\subsection{Tuning process}

In order to invoke lambda function concurrently, Seneca integrates with Celery~\cite{ref:celery}. Celery is an asynchronous task queue based on distributed message passing. Celery workers are working processes that pick up tasks from the queue, execute tasks and store the asynchronous result at backend (Redis~\cite{ref:redis} in our case). Based on user-defined configuration file, Seneca creates a list of payloads that contains values of hyperparameters and the whole list covers the entire Cartesian product search space. Every payload launches a Celery task, on which a worker executes by triggering lambda function and insert a new record of the returned evaluation score and corresponding hyperparameter setting to backend database.

After the comprehensive searching, Seneca will extract all records within search space and select the best result. Up to users' specification, Seneca returns either hyperparameter setting or a serialized model trained by this setting.

### 2019-02-02 results.tex
mz - Add background, Seneca, related_work

For these applications (XGBoost, SVC, and Neural Network), we consider three
split scenarios for training and testing: 80\%-20\%, 20\%-80\% and
0.1\%-99.9\%. As the proportion of training data decreases, the performance of
classifier degrades accordingly.  However we note that using the optimized
hyperparameters generated by Seneca, these calssifiers are able to achieve
accuracy levels similar to those generated from the default parameters, but
with significantly less training data.

@@

To evaluate the cost savings introduced by Seneca,
it on Amazon Web Services (AWS) Lambda~\cite{XXXaws-lambdaXXX} and extract
execution time and memory use from AWS CloudWatch~\cite{XXXcloudwatchCXXX} logs.  
We then compute monetary cost
using the AWS Lambda pricing model~\cite{ref:pricing}.
%awslambdapricing: https://aws.amazon.com/lambda/pricing/
Each function downloads the training and testing dataset 
of the application from AWS Simple Storage Service (S3) upon function invocation. 
We do not consider the cost of dataset storage in our cost computations (it is very small).

@@

\begin{figure}[t] \centering 
\includegraphics[scale=0.36]{box_plot_accuracy}
\caption{The box plot of accuracy metric from three classification applications for the entire hyperparameter tuning search space. The red notch indicates the accuracy metric from the trained model using default settings. The colorful diamonds are outliers beyond two interquartile ranges. Higher accuracy values are better.
\label{fig:box_plot_accuracy}}
% \vspace{-0.2in}
\end{figure}

@@

These tables illustrate two important cost features of Seneca.  First, using
AWS Lambda, full hyperparameter space exploration is not expensive in 
\textit{absolute} dollar cost terms.  Even using the maximum memory available
for almost 25 minutes of execution time (cf Prophet in Table~\ref{XXX}), the
total dollar cost is approximately 73 cents.  Secondly, Seneca's memory
optimizations produce a significant percentage savings which, over repeated
executions, is important.

Note also that Seneca's memory/cost optimzation
introduces at an execution time penalty (relative to the maximum memory
execution) for some of the applications.  That is, the
execution times for all applications except SVC increase when Seneca chooses
a memory size.  Thus while Seneca is finding the best application output
quality and optimizing the cost of producing that output, it is doing so at
the expense of execution performance (time to solution).    
The source of this execution performance penalty is not clear.  We hypothesize
that when requesting a large memory for AWS Lambda, Amazon chooses a machine
that also has fast processors so the Lambda functions execute more quickly.
If this supposition is true, then the AWS Lambda pricing model correctly
weights the cost of additional memory (in light of the reduced execution time)
because the large-memory executions are still more expensive.
However regardless of the cause, the Seneca memory optimizer 
provides a way to trade off time to solution for cost when the highest
possible application output quality is desired.
% \textcolor{blue}{See my notes to fix the tables in the caption of these figures.}

>>>>

% For these applications (XGBoost, SVC, and Neural Network), we consider three
% split scenarios for training and testing: 80\%-20\%, 20\%-80\% and
% 0.1\%-99.9\%. As the proportion of training data decreases, the performance of
% classifier degrades accordingly. However we note that using the optimized
% hyperparameters generated by Seneca, these classifiers are able to achieve
% accuracy levels similar to those generated from the default parameters, but
% with significantly less training data.

@@

To evaluate the cost savings introduced by Seneca,
it on Amazon Web Services (AWS) Lambda~\cite{ref:awslambda} and extract
execution time and memory use from AWS CloudWatch~\cite{ref:awscloudwatch} logs.  
We then compute monetary cost
using the AWS Lambda pricing model~\cite{ref:pricing}.
%awslambdapricing: https://aws.amazon.com/lambda/pricing/
Each function downloads the training and testing dataset 
of the application from AWS Simple Storage Service (S3)~\cite{ref:awss3} upon function invocation. 
We do not consider the cost of dataset storage in our cost computations (it is very small).

@@

\begin{figure}[t] \centering 
\vspace{-0.2in}
\includegraphics[scale=0.36]{box_plot_Accuracy_20}
\caption{The box plot of accuracy metric from three classification applications for the entire hyperparameter tuning search space. The red notch indicates the accuracy metric from the trained model using default settings. The colorful diamonds are outliers beyond two interquartile ranges. Higher accuracy values are better.
\label{fig:box_plot_accuracy}}
\vspace{-0.1in}
\end{figure}

@@

These tables illustrate two important cost features of Seneca.  First, using
AWS Lambda, full hyperparameter space exploration is not expensive in 
\textit{absolute} dollar cost terms.  Even using the maximum memory available
for almost 25 minutes of execution time (cf Prophet in Table~\ref{tab:cost_max_memory}), the
total dollar cost is approximately 73 cents.  Secondly, Seneca's memory
optimizations produce a significant percentage savings which, over repeated
executions, is important.

Note also that Seneca's memory/cost optimization
introduces at an execution time penalty (relative to the maximum memory
execution) for some of the applications.  That is, the
execution times for all applications except SVC increase when Seneca chooses
a memory size.  Thus while Seneca is finding the best application output
quality and optimizing the cost of producing that output, it is doing so at
the expense of execution performance (time to solution). Such penalty is expected due to the proportional CPU power to allocated memory. However, Seneca optimizer is able to find the optimal memory configuration to minimize such penalty. In the case of SVC, Seneca saves 30\% compared to max memory execution (\$0.016 vs \$0.023) but has the same execution time (0.91 minutes).
% The source of this execution performance penalty is not clear.  We hypothesize
% that when requesting a large memory for AWS Lambda, Amazon chooses a machine
% that also has fast processors so the Lambda functions execute more quickly.
% If this supposition is true, then the AWS Lambda pricing model correctly
% weights the cost of additional memory (in light of the reduced execution time)
% because the large-memory executions are still more expensive.
%However regardless of the cause, 
Even in the worst case of CPU-bound application, the Seneca memory optimizer still provides a way to trade off time to solution for cost when the highest possible application output quality is desired.
% \textcolor{blue}{AWS announces the proportional CPU power to memory which is the source of penalty}
% \textcolor{blue}{See my notes to fix the tables in the caption of these figures.}
